### 2016-06-22 aosd13/abstract.tex

\begin{abstract}
Modularity and efficiency are often contradicting requirements, such that programers have
to trade one for the other. We analyze this dilemma in the context of programs
operating on collections. Performance-critical code using collections need often to be hand-optimized, leading to non-modular, brittle, 
and redundant code. In principle, this dilemma could be avoided by automatic collection-specific
optimizations, such as fusion of collection traversals, usage of indexing, or
reordering of filters. Unfortunately, it is not obvious how to encode such optimizations 
in terms of ordinary collection APIs, because the program operating on the collections is not reified and hence cannot be analyzed.

We propose \LoS, the Scala Query Optimizer---a \emph{deep embedding} of the Scala collections API
that allows such analyses and optimizations to be defined and executed within Scala, without relying on external tools 
or compiler extensions. \LoS\ provides the same ``look and feel'' (syntax and
static typing guarantees) as the standard collections API\@.
We evaluate \LoS\ by re-implementing several code analyses of the FindBugs tool
using \LoS{}, show average speedups of \avgSpeedupT{} with a maximum of
\maxSpeedupT{} and hence demonstrate that \LoS\ can
reconcile modularity and efficiency in real-world applications.
\end{abstract}

%\category{CR-number}{subcategory}{third-level}
\category{H.2.3}{Database Management}{Languages}[Query languages]
%\category{D.3.3}{Programming Languages}{Language Constructs and Features}
\category{D.1.1}{Programming Techniques}{Applicative (Functional) Programming}
\category{D.1.5}{Programming Techniques}{Object-oriented Programming}

% Not required on first page.
%\terms
%Design, Performance, Languages

\keywords
Deep embedding; query languages; optimization; modularity