### 2016-06-22 aosd13/graphs/newTable.tex

$1$&{\tablerowsize Covariant compareTo() defined}&1.1&1.3&0.85&0.26&0.26&0.9&5.0&4.4\\
$2$&{\tablerowsize Explicit garbage collection call}&496&258&1176&1150&52&1.9&0.2&9.5\\
$3$&{\tablerowsize Protected field in final class}&11&1.1&11&1.2&1.2&10.0&1.0&9.8\\
$4$&{\tablerowsize Explicit runFinalizersOnExit() call}&509&262&1150&1123&10.0&1.9&0.2&51\hphantom{.0}\\
$5$&{\tablerowsize clone() defined in non-Cloneable class}&29&14&55&46&0.47&2.1&0.3&61\hphantom{.0}\\
$6$&{\tablerowsize Covariant equals() defined}&29&15&23&9.7&0.20&1.9&1.6&147\hphantom{.0}\\
$7$&{\tablerowsize Public finalizer defined}&29&12&28&8.0&0.03&2.3&1.5&1070\hphantom{.0}\\
$8$&{\tablerowsize Dubious catching of IllegalMonitorStateException}&82&72&110&28&0.01&1.1&2.6&12800\hphantom{.0}\\
$9^R$&{\tablerowsize Uninit. field read during construction of super}&896&367&3017&960&960&2.4&0.4&0.9\\
$10^R$&{\tablerowsize Mutable static field declared public}&9527&9511&9115&9350&9350&1.0&1.0&1.0\\
$11^R$&{\tablerowsize Refactor anon. inner class to static}&8804&8767&8718&8700&8700&1.0&1.0&1.0\\
$12^R$&{\tablerowsize Inefficient use of toArray(Object[])}&3714&1905&4046&3414&3414&2.0&0.6&1.1\\
$13^R$&{\tablerowsize Primitive boxed and unboxed for coercion}&3905&1672&5044&3224&3224&2.3&0.5&1.2\\
$14^R$&{\tablerowsize Double precision conversion from 32 bit}&3887&1796&5289&3010&3010&2.2&0.6&1.3\\
$15^R$&{\tablerowsize Privileged method used outside doPrivileged}&505&302&1319&337&337&1.7&0.9&1.5\\
$16^R$&{\tablerowsize Mutable public static field should be final}&13&6.2&12&7.0&7.0&2.0&0.9&1.8\\
$17^R$&{\tablerowsize Serializable class is member of non-ser. class}&12&0.77&0.94&1.8&1.8&16\hphantom{.0}&0.4&6.9\\
$18^R$&{\tablerowsize Swing methods used outside Swing thread}&577&53&1163&45&45&11\hphantom{.0}&1.2&13\hphantom{.0}\\
$19^R$&{\tablerowsize Finalizer only calls super class finalize}&55&13&73&11&0.10&4.4&1.1&541\hphantom{.0}\\