### 2016-06-22 aosd13/graphs/table.tex

{\tablerowsize Covariant compareTo() defined}&$1.29$&$\pm$&$0.18$&$1.14$&$\pm$&$0.02$&$29.0$&$\pm$&$0.84$&1.2x&$ 2.2 $&$\pm$&$ 0.3 $x&$ 0 $&$\pm$&$ 0 $x\\
{\tablerowsize Explicit garbage collection call}&$257.53$&$\pm$&$9.92$&$496.35$&$\pm$&$7.17$&$52.0$&$\pm$&$1.66$&0.2x&$ 4.9 $&$\pm$&$ 0.4 $x&$ 2.5 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Protected field in final class}&$1.13$&$\pm$&$0.02$&$11.31$&$\pm$&$0.58$&$3.5$&$\pm$&$0.14$&0.8x&$ 1 $&$\pm$&$ 0 $x&$ 0.2 $&$\pm$&$ 0 $x\\
{\tablerowsize Explicit runFinalizersOnExit() call}&$261.67$&$\pm$&$8.08$&$509.44$&$\pm$&$6.44$&$12.0$&$\pm$&$0.44$&0.2x&$ 26.2 $&$\pm$&$ 2.9 $x&$ 12 $&$\pm$&$ 0.7 $x\\
{\tablerowsize clone() defined in non-Cloneable class}&$13.92$&$\pm$&$0.24$&$28.66$&$\pm$&$0.83$&$13.0$&$\pm$&$0.45$&0.3x&$ 29.7 $&$\pm$&$ 3.1 $x&$ 1 $&$\pm$&$ 0 $x\\
{\tablerowsize Covariant equals() defined}&$15.23$&$\pm$&$0.35$&$29.29$&$\pm$&$0.87$&$15.0$&$\pm$&$0.36$&1.6x&$ 38.5 $&$\pm$&$ 23.5 $x&$ 1 $&$\pm$&$ 0 $x\\
{\tablerowsize Public finalizer defined}&$12.20$&$\pm$&$0.20$&$28.68$&$\pm$&$0.81$&$7.5$&$\pm$&$0.24$&1.0x&$ 457.1 $&$\pm$&$ 53 $x&$ 1.6 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Dubious catching of IllegalMonitorStateException}&$71.72$&$\pm$&$7.83$&$82.42$&$\pm$&$2.62$&$9.4$&$\pm$&$0.34$&1.0x&$ 11142.7 $&$\pm$&$ 1598.1 $x&$ 7.6 $&$\pm$&$ 0.9 $x\\
{\tablerowsize Uninit. field read during construction of super}&$366.55$&$\pm$&$6.78$&$895.97$&$\pm$&$32.91$&$380.0$&$\pm$&$10.90$&0.2x&$ 0.4 $&$\pm$&$ 0 $x&$ 0.3 $&$\pm$&$ 0 $x\\
{\tablerowsize Mutable static field declared public}&$9510.84$&$\pm$&$523.38$&$9526.54$&$\pm$&$574.77$&$91.0$&$\pm$&$2.74$&1.0x&$ 1 $&$\pm$&$ 0.1 $x&$ 1 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Refactor anon. inner class to static}&$8766.79$&$\pm$&$527.92$&$8803.56$&$\pm$&$538.20$&$120.0$&$\pm$&$3.67$&1.0x&$ 1 $&$\pm$&$ 0.1 $x&$ 1 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Inefficient use of toArray(Object[])}&$1904.82$&$\pm$&$106.96$&$3714.45$&$\pm$&$155.22$&$89.0$&$\pm$&$2.51$&0.8x&$ 0.8 $&$\pm$&$ 0.1 $x&$ 0.8 $&$\pm$&$ 0 $x\\
{\tablerowsize Primitive boxed and unboxed for coercion}&$1671.65$&$\pm$&$118.27$&$3905.43$&$\pm$&$301.70$&$59.0$&$\pm$&$1.44$&0.8x&$ 0.9 $&$\pm$&$ 0.1 $x&$ 0.9 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Double precision conversion from 32 bit}&$1795.66$&$\pm$&$102.41$&$3887.10$&$\pm$&$302.64$&$40.0$&$\pm$&$1.22$&0.7x&$ 0.8 $&$\pm$&$ 0.1 $x&$ 0.8 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Privileged method used outside doPrivileged}&$301.68$&$\pm$&$32.02$&$505.32$&$\pm$&$30.56$&$27.0$&$\pm$&$0.84$&0.2x&$ 1 $&$\pm$&$ 0.1 $x&$ 0.9 $&$\pm$&$ 0.1 $x\\
{\tablerowsize Mutable public static field should be final}&$6.16$&$\pm$&$0.20$&$12.63$&$\pm$&$0.59$&$15.0$&$\pm$&$0.49$&0.8x&$ 1 $&$\pm$&$ 0 $x&$ 0.3 $&$\pm$&$ 0 $x\\
{\tablerowsize Serializable class is member of non-ser. class}&$0.77$&$\pm$&$0.38$&$12.14$&$\pm$&$0.64$&$9.5$&$\pm$&$0.32$&0.5x&$ 0.4 $&$\pm$&$ 0.4 $x&$ 0.1 $&$\pm$&$ 0 $x\\
{\tablerowsize Swing methods used outside Swing thread}&$53.37$&$\pm$&$0.91$&$577.22$&$\pm$&$33.65$&$46.0$&$\pm$&$1.49$&0.9x&$ 1 $&$\pm$&$ 0 $x&$ 0.5 $&$\pm$&$ 0 $x\\
{\tablerowsize Finalizer only calls super class finalize}&$12.53$&$\pm$&$0.21$&$55.42$&$\pm$&$1.29$&$26.0$&$\pm$&$0.63$&0.5x&$ 73.2 $&$\pm$&$ 22.5 $x&$ 0.5 $&$\pm$&$ 0 $x\\