### 2016-06-22 aosd13/pepm13/abstract.tex

\begin{abstract}
The collections API of a programming language forms an embedded domain-specific language to express queries and operations
on collections. Unfortunately, the ordinary style of implementing such APIs does not allow automatic domain-specific
analyses and optimizations such as fusion of collection traversals, usage of indexing, or reordering of filters.
Performance-critical code using collections must instead be hand-optimized, leading to non-modular, brittle, 
and redundant code.

We propose \LoS, the Scala Query Optimizer---a \emph{deep embedding} of the Scala collections API
that allows such analyses and optimizations to be defined and executed within Scala, without relying on external tools 
or compiler extensions. \LoS\ provides the same ``look and feel'' (syntax and static typing guarantees) as the standard collections API.
We evaluate \LoS\ by re-implementing several code analyses of the Findbugs tool using \LoS{} and demonstrate that \LoS\ can
reconcile modularity and efficiency in real-world applications.

Overall, we present the collections APIs as a critical case study
for DSEL-implementation techniques. From it, we extract lessons about the design of embedded DSLs and host-language design for 
effective and type-safe embedding and optimizations.

%% Trying to follow Kent Beck's idea (http://www.sigplan.org/oopsla/oopsla96/how93.html).
%% Domain-specific abstractions are not part of general-purpose languages. Such abstractions are important to close the abstraction gap between a solution to a problem and its encoding as a program. Domain-specific embedded languages support such abstractions but are only part of the solution: support for meta-programming on the embedded language is also essential, to support more powerful abstractions and to allow domain-specific optimizations.
%%we argue that reifying programs in the embedded language is to support metaprogramming and domain-specific optimizations can be crucial.
%% To investigate support for this requirements, we consider Scala, a language intended to support DSEL engineering, and present a case study on a query DSL for collections: without changes to the compiler, we reify queries on Scala collections while preserving typing and show how to implement algebraic optimizations to provide order-of-magnitude speed-ups and make declarative queries efficient; finally we extract lessons about how the host language supports the development of our DSEL.
%% Finally we discuss lessons for DSEL developers and for programming languages.
%
%%Deeply embedded domain-specific languages are a worthwhile approach
%%Domain-specific embedded languages
%
%%Queries over collections need to be efficient, and to this end they are often optimized by hand instead of
%%being specified in a declarative way. Database optimizers free the developer from manual optimizations of queries
%%over databases, but are often not applied to collections. Employing the same techniques for queries over collections
%%allow to improve performance, to make queries more compact and readable, or both.
%%
%%We implement a deeply-embedded DSEL for queries over collections in Scala, close to native query operators,
%%implement a few significant optimizations, and demonstrate their usefulness over bug-finding queries on source code
%%extracted from FindBugs.
\end{abstract}

\begin{oopsla}
% Klaus: Let's forget about these things until the paper gets accepted
%\category{CR-number}{subcategory}{third-level}

\category{H.2.3}{Database Management}{Languages}[Query languages]
%\category{D.3.3}{Programming Languages}{Language Constructs and Features}
\category{D.1.1}{Programming Techniques}{Applicative (Functional) Programming}
\category{D.1.5}{Programming Techniques}{Object-oriented Programming}

\terms
Design, Performance, Languages

\keywords
%keyword1, keyword2
Deep embedding, query languages, optimization, modularity
\end{oopsla}