### 2016-06-22 aosd13/aosd13-extended-appendixes.tex
aosd13: adjust relative paths

\section{Collections as a case study}
\label{sec:caseStudy}
As discussed in the introduction, to support optimizations we require a deep embedding of the collections DSL\@.
\input{aosd13/pepm13/caseStudyGoals}

\section{Implementation: Expressing the interface in Scala}
\label{sec:intfScala}
\input{aosd13/pepm13/caseStudyImpl}

\section{Discussion}
\label{sec:discussion}
\input{aosd13/pepm13/caseStudyDiscussion}
\pg{TODO: mention discussion with Jacques Carette. Maybe mention that in the technical report only?}
\input{aosd13/pepm13/caseStudyRW}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: