### 2016-06-22 main.tex
Create new main file
Some of this code comes from my PLDI14 paper.

% Ensure that arXiv processes this as PDF.
\pdfoutput=1

\documentclass{book}
\usepackage{natbib}
\usepackage{comment}
\bibliographystyle{abbrvnat}
\input{packages}
\input{macros}

\usepackage{amsthm}
\usepackage{listings,fixfoot}
\input{aosd13/lstsetup}

\input{aosd13/customMacros}
%\input{pldi14/macros}
%\input{macros_part2}
% from pldi14
\newcommand{\ILC}{ILC}
% from aosd13
\usepackage{comment}
\excludecomment{extraEval}
\includecomment{techrep}
\excludecomment{nontechrep}
\excludecomment{todos}
\includecomment{forNonBeamer}

\title{PhD Thesis}
\author{Paolo G. Giarrusso}
\begin{document}
\maketitle
\chapter{Intro}
\chapter{Reifying collection queries}

\input{aosd13/graphs/evalResLos}
\input{aosd13/paperBody}
\input{aosd13/aosd13-extended-appendixes}

\bibliography{Bibs/DB,Bibs/ProgLang,Bibs/SoftEng,Bibs/own}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: