### 2016-06-22 pldi14/fig-function-change.tex
Adds pldi14 paper
Selected files from ps-mr/ilc@233142c7f93f2e9db25d9b540af370ce6cbc28c8.

% Emacs, this is -*- latex -*-!
\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{y=2cm}
\path
(-90:1)node(fv)  {$\App {\Old f} {\Old x}$}
( 90:1)node(f'v'){$\App{\Apply*{\D f} {\Old f}} {\Apply*{\D x} {\Old x}}$}
;
\draw[-stealth](fv)--(f'v')
node[pos=.5,right]{$\App{\App{\D f}{\Old x}}{\D x}$};
\end{tikzpicture}
\caption{Functional changes react to input changes and produce
the difference between the new function on the new input and the
old function on the old input:
$\GD(\Gs\r\Gt)=\Gs\r\GD\Gs\r\GD\Gt$.}
\label{fig:function-change}
\end{figure}