### 2016-06-22 pldi14/sec-abstract.tex

% Emacs, this is -*- latex -*-!
If the result of an expensive computation is invalidated by a small
change to the input, the old result
should be updated incrementally instead of reexecuting the whole computation.
We incrementalize programs through
their \emph{derivative}. A derivative maps
changes in the program's input directly to changes in the program's
output, without reexecuting the original program. We present a
program transformation taking programs to their
derivatives, which is fully static and automatic, supports first-class
functions, and produces derivatives amenable to standard
optimization.

We prove the program transformation correct in Agda for a
family of simply-typed $\lambda$-calculi, parameterized by base
types and primitives. A precise interface specifies what is
required to incrementalize the chosen primitives.

We investigate performance by a case study: We implement in Scala the
program transformation, a plugin and improve performance of a nontrivial program by
orders of magnitude.

% \pg{Failed attempt at four-sentence abstract. Our paper isn't that cool yet. Hmpf!}
% \pg{I submitted what's above for now. But maybe we should still try to have
% something like this?}
% Results of expensive computations are often recomputed from scratch when
% the input changes. Often, this wastes time because changes are
% small and the output could be updated much faster; writing
% code performing efficient updates by hand is error-prone. We propose
% a program transformation taking a program to its
% \emph{derivative}, mapping changes to inputs to changes to
% outputs, often more efficiently, and we perform case studies
% showing significant performance improvements.