### 2016-06-22 pldi14/sec-concl.tex

\section{Conclusions and future work}
\label{sec:concl}
\label{ssec:future}
We have presented \ILC, an approach to lifting incremental computations
on first-order programs to incremental computations on higher-order
programs. We have presented a machine-checked correctness proof 
of a formalization of \ILC\ and an initial experimental evaluation
in the form of an implementation, a sample plugin for maps and bags,
and a non-trivial example that was incrementalized successfully and
efficiently.

Our work opens several avenues of future work. Our current implementation
is not efficient on derivatives that are not self-maintainable.
However, as discussed
(Sec.~\ref{ssec:self-maint}), we will study how
to memoize intermediate results to address this limitation. Our next
step will be to develop language plugins which
have efficient non-self-maintainable primitives.

Another area of future work is adding support for algebraic data
types (including recursive types), polymorphism, subtyping, general recursion
and other collection types. While support for algebraic data
types could subsume support for specific collections, many
collections have additional algebraic properties that enable faster
incrementalization (like bags). Even lists (which have fewer algebraic properties)
can benefit from special support~\citep{Maier2013}.

Moreover, we intend to apply \ILC{} to optimize queries on
collections in the context of the \textsc{SQuOpt}
project~\citep{GiarrussoAOSD13}, which was a motivation for this
work; in particular, \textsc{SQuOpt} can automatically rewrite
queries to use database-style indexes, and \ILC{} enables
updating those indexes when input data changes.

\begin{oldSec}
Our derivatives accept arbitrary changes for all inputs. We plan
to stage the code that updates the inputs through lightweight
modular staging~\citep{rompf2010lightweight}, analyze it to
predict changes and specialize the derivative to the result of
this analysis. For instance, if we need to update $\Term{f} \;
\Term{bag}$ after $\Term{bagNew} = \Term{insert} \; \Term{el} \;
\Term{bagOld}$, we can infer that $\Term{dbag}$ is an insertion
and specialize $f'$ to such a change.
\end{oldSec}

Finally, we intend to perform a full and thorough experimental evaluation
to demonstrate that \ILC\ can incrementalize large-scale practical programs.