### 2016-06-22 pldi14/sec-formal.tex

% Emacs, this is -*- latex -*-!
\section{Notes on mechanized formalization}
\label{sec:formal}

The mechanization of our proof uses the Agda
proof assistant~\citep{agda-head}.

Agda is an implementation of intensional Martin-Löf type theory.
For our proofs, we need to postulate that equality of functions is
extensional; this postulate is known to be consistent with Agda's
type theory~\citep{Hofmann96}, hence it is safe to assume in Agda%
\footnote{\url{http://permalink.gmane.org/gmane.comp.lang.agda/2343}}.

Moreover, we postulate a few standard axioms on the
implementation of bags, to avoid proving correct an
implementation of bags, or needing to account for different
values representing the same bag (such different values typically
arise when implementing bags as search tree).

We formalize the simply-typed lambda calculus we presented using
typed de Brujin indexes to handle binding, because it takes well
advantage of Agda's support for type refinement in pattern
matching. On top of that, we implement a HOAS-like frontend, that
we use for writing specific terms.
The domains for our denotational
semantics are Agda types, and semantic values are simply Agda
values---in other words, we give a denotational semantics in
terms of type theory.
%
This allows us to state the specification of differentiation
directly in the semantic domain, and take advantage of Agda's
support for equational reasoning between Agda functions.

The main technical difference with the presentation we have given
is that we actually use dependent types to define change
structures: $\Change{v}$ is the \emph{type} of changes valid for
$v$. Moreover, we work in proof-relevant mathematics, so that for
instance $\Change[\Fun*{\Gs}{\Gt}]{f}$ is a $\Sigma$-type, that
is a dependent pair containing a value and a proof.