### 2016-08-05 main.tex
Start sketching new chapters

% Ensure that arXiv processes this as PDF.
\pdfoutput=1

\documentclass{book}
\usepackage{natbib}
\usepackage{comment}
\bibliographystyle{abbrvnat}
\input{packages}
\input{macros}

\usepackage{amsthm}
\usepackage{listings,fixfoot}
\input{lstsetup}

\input{aosd13/customMacros}
\input{pldi14/customMacros}

\input{pldi14/macros_part2}
% from pldi14
\newcommand{\ILC}{ILC}
% from aosd13
\usepackage{comment}
\excludecomment{extraEval}
\includecomment{techrep}
\excludecomment{nontechrep}
\excludecomment{todos}
\includecomment{forNonBeamer}

\title{PhD Thesis}
\author{Paolo G. Giarrusso}
\begin{document}
\maketitle
\chapter{Intro}
\chapter{Reifying collection queries}

\input{aosd13/graphs/evalResLos}
\input{aosd13/paperBody}
\input{aosd13/aosd13-extended-appendixes}

\newcommand{\co}[1]{\code{#1}} %Adaptation between different interfaces
\chapter{Incrementalizing simply-typed \TitleLambda{}-calculus}
\input{pldi14/agda}

\input{pldi14/sec-intro}
\input{pldi14/sec-change-theory}
\input{pldi14/sec-differentiate}
\input{pldi14/sec-practice}
\input{pldi14/sec-rw}
\input{pldi14/sec-concl}

% Appendixes
\input{pldi14/sec-evaluation}
\input{pldi14/sec-addendum}

\input{pldi14/sec-formal}

\input{pldi14/sec-change-structures}

\begin{oldSec}
\input{pldi14/sec-STLC-correct}
\end{oldSec}

\begin{oldSec}
\input{pldi14/sec-informal}
\end{oldSec}

\chapter{Incrementalizing more programs}
\pg{Only a sketch for now}

While the basic framework we presented is significant, applying it in practice
to incrementalize programs is hard because of a few problems.
\begin{enumerate}
\item The incremental computation does not have accesss to intermediate results from the original computation.
\item Since function changes are represented as functions, the derivative cannot
  test if a function change is nil. In fact, while a function change can replace
  a function with an arbitrary other function, actual changes often simply
  replace a closure with another closure using the same code but closing over a
  changed environment.
\item Applying a derivative to a nil change always produce a nil change, but we
  never take advantage of this to optimize derivatives, except sometimes at
  compile time.
\item We do not support change composition.
\end{enumerate}

\bibliography{Bibs/DB,Bibs/ProgLang,Bibs/SoftEng,Bibs/own}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: