### 2016-12-03 pldi14/macros_part2.tex
Number definitions correctly
This adapts the PLDI setup to a chapter-based document.

% The next package must be loaded last.
% It adds the \cref{label} command, producing Eq./Sec./whatever followed by the reference.
\usepackage[capitalise]{cleveref}

\crefname{section}{Sec.}{Sec.}
\crefname{subsection}{Sec.}{Sec.}
\crefname{subsubsection}{Sec.}{Sec.}

% Theorems, lemmas, corollaries
% Typeset in roman to avoid known TeX feature
% (bad spacing between italic text and math formulas)
%
% Numbering:
% Theorem 3.1. It always rains.
% Lemma 3.2. If it rains, then the weather is bad.
% Corollary 3.3. The weather is always bad.
%
% http://tex.stackexchange.com/a/67251
\theoremstyle{definition}
\input{pldi14/QED}
\newtheorem{theorem}{Theorem}
% http://tex.stackexchange.com/a/45358/1340
% \numberwithin{section}{chapter} % From AMS Author FAQ (http://www.ams.org/faq?class_id=561), makes no difference?
\numberwithin{theorem}{section}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}

\begin{oldSec}
\newtheorem{parameter}[theorem]{Plugin Requirement}
\end{oldSec}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Enumerations inside theorem environments
%
% (cannot be in macros.tex because of \crefname)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newlist{subdefinition}{enumerate}{1}
\setlist*[subdefinition]{label=(\alph*), ref=\arabic{chapter}.\arabic{section}.\arabic{theorem}\alph*}
\crefname{subdefinitioni}{Property}{Properties}
\Crefname{subdefinitioni}{Property}{Properties}

\newlist{subparameter}{enumerate}{1}
\setlist*[subparameter]{label=(\alph*), ref=\arabic{chapter}.\arabic{section}.\arabic{theorem}\alph*}
\crefname{subparameteri}{Plugin Requirement}{Plugin Requirements}
\Crefname{subparameteri}{Plugin Requirement}{Plugin Requirements}

\newlist{subtheorem}{enumerate}{1}
\setlist*[subtheorem]{label=(\alph*), ref=\arabic{chapter}.\arabic{section}.\arabic{theorem}\alph*}
\crefname{subtheoremi}{Theorem}{Theorems}
\Crefname{subtheoremi}{Theorem}{Theorems}