### 2016-12-20 aosd13/aosd13-extended-appendixes.tex
Turn AOSD'13 PEPM appendixes into own chapter

\chapter{Collections as a case study}
\label{sec:caseStudy}
In this chapter we discuss collections as a \emph{critical} case
study~\citep{flyvbjerg06five} for deep DSL embedding.

As discussed in the previous chapter, to support optimizations we require a deep embedding of the collections DSL\@.

\input{aosd13/pepm13/caseStudyGoals}

\section{Implementation: Expressing the interface in Scala}
\label{sec:intfScala}
\input{aosd13/pepm13/caseStudyImpl}

\section{Discussion}
\label{sec:discussion}
\input{aosd13/pepm13/caseStudyDiscussion}
\pg{TODO: mention discussion with Jacques Carette. Maybe mention that in the technical report only?}
\input{aosd13/pepm13/caseStudyRW}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: