### 2017-07-19 Abstract.tex
Create separate file Abstract.tex

\section*{Abstract}
 Programs manipulating data collections require

% \begin{abstract}
% \end{abstract}