### 2017-07-19 popl18/static-caching.tex
static-caching.tex: move more code into preamble

%&static-caching
% The above line loads a precompiled format coming from precompiling static-caching.ltx.

% Prelude code which changes more often should appear below.

\begin{document}
\maketitle

\newcommand{\poplPath}[1]{#1}
\input{cts-intro}
\input{cts-motivation}
\input{formalization}
\input{cts-case-studies}
\input{cts-relwork}
\input{cts-conclusion}

\bibliography{../Bibs/ProgLang,../Bibs/DB,../Bibs/own,../Bibs/SoftEng}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: