### 2017-07-21 popl18/static-caching.tex
Popl18: externalize tikz pictures, save 1sec (30% build time!)

%&static-caching
% The above line loads a precompiled format coming from precompiling the "static
% preamble" static-caching.ltx.

% Prelude code which changes more often should appear below.

% Build tikz pictures into separate files. Relies on setup in static preamble.
\tikzexternalize
\begin{document}
\maketitle

\newcommand{\poplPath}[1]{#1}
\input{cts-intro}
\input{cts-motivation}
\input{formalization}
\input{cts-case-studies}
\input{cts-relwork}
\input{cts-conclusion}

\bibliography{../Bibs/ProgLang,../Bibs/DB,../Bibs/own,../Bibs/SoftEng}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: