### 2017-07-21 acks.tex
Split acks out to reorder them
And add addcontentsline rule for Acknowledgments, like for Bibliography, so
they're in the TOC and bookmarks.

% Emacs, this is -*- latex -*-!

\chapter*{Acknowledgments}
\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgments}
% From AOSD13
I thank Sebastian Erdweg for helpful discussions on
this project, Katharina Haselhorst for help
implementing the code generator, and the anonymous reviewers, Jacques Carette and Karl Klose
for their helpful comments on this chapter.
This work is supported in part by the European Research Council, grant \#203099 ``ScalPL''.

% From PLDI14 (?)
% From ILC17
We thank Cai Yufei, Tillmann Rendel, Lourdes Del Carmen Gonz\`alez Huesca, Yann
R\`egis-Gianas, Philipp Schuster, Sebastian Erdweg, Marc Lasson, Robert Atkey,
... for helpful discussions on this project.