### 2017-07-21 Abstract.tex
Mention all lists, biblios in ToC/bookmarks, reorder

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

 Programs manipulating data collections require

% \begin{abstract}
% \end{abstract}