### 2017-07-21 Abstract.tex
Simplify last commit via tocbibind and \frontmatter

% Thanks to \frontmatter, this is unnumbered.
% https://tex.stackexchange.com/a/244746/1340
\chapter{Abstract}

 Programs manipulating data collections require

% \begin{abstract}
% \end{abstract}