### 2017-07-22 acks.tex
Better acks

% Emacs, this is -*- latex -*-!

\chapter{Acknowledgments}
I thank my supervisor Klaus Ostermann for encouraging me to do a PhD
and for his years of supervision. He has always been patient and kind.

Thanks also to Fritz Henglein for agreeing to review my thesis.

I have many people to thank and too little time to do it properly: I'll finish
these acknowledgments in the final version.

The research presented here has benefited from the help of coauthors,
reviewers and people who offered helpful comments.
Moreover, this PhD thesis wouldn't be possible without all the people who have
supported me, both during all the years of my PhD, but also in the years before.

% From AOSD13

% I thank Sebastian Erdweg for helpful discussions on
% this project, Katharina Haselhorst for help
% implementing the code generator, and the anonymous reviewers, Jacques Carette and Karl Klose
% for their helpful comments on this chapter.

% % From PLDI14 (?)
% % From ILC17
% I thank Cai Yufei, Tillmann Rendel, Lourdes Del Carmen Gonz\`alez Huesca, Yann
% R\`egis-Gianas, Philipp Schuster, Sebastian Erdweg, Robert Atkey, Gabriel
% Scherer and Marc Lasson for helpful discussions on this project.