### 2017-08-02 TitlePage.tex
Switch to lmodern for whole thesis
Output seems close enough (just a couple page breaks shifted). Chapter title
fonts are bolder.

\begin{titlepage}
% Declarations, in scope until \end{titlepage}.
\makeatletter % for \@title and \@author
\setlength{\parskip}{0pt}
\centering
\large
%
% https://tex.stackexchange.com/a/241121/1340
\linespread{1.5}
\sffamily
% Slanted because no real \itshape available for this font
% (https://tex.stackexchange.com/a/68932/1340).
\slshape
% Huge titles are broken by cm-super. Using lmodern fixes it.
\selectfont

\vspace*{-2cm}

{\huge \bfseries
  \@title%
  \par% Must be in scope of the \huge declaration to get correct spacing.
}

\vspace{4cm}

{\bfseries Dissertation}\\
der Mathematisch- und Naturwissenschaftlichen Fakult\"at\\
der Eberhard Karls Universit\"at T\"ubingen\\
zur Erlangung des Grades eines\\
Doktors der Naturwissenschaften\\
(Dr.\ rer.\ nat.)

\vspace{4 cm}

vorgelegt von \\
\@author \\
aus Catania (Italien)

\vspace{2 cm}

T\"ubingen\\
2018

\vfill
\end{titlepage}
\clearpage
\thispagestyle{empty}
%
\vspace*{\fill}

Gedruckt mit Genehmigung der Mathematisch-Naturwissenschaftlichen Fakultät der
Eberhard Karls Universität Tübingen.

% 3 bigskips in a row.
% https://tex.stackexchange.com/a/245663/1340
\vspace{\dimexpr \bigskipamount * 3 \relax}
%
% https://tex.stackexchange.com/a/16261/1340
\begin{tabular}{@{}ll@{}}
Tag der mündlichen Qualifikation: & 26.01.2018\\
Dekan:               & Prof.\ Dr.\ Wolfgang Rosenstiel\\
1.\ Berichterstatter: & Prof.\ Dr.\ Klaus Ostermann\\
2.\ Berichterstatter: & Prof.\ Dr.\ Fritz Henglein
\end{tabular}
%
\vspace{-1cm} %XXX revise when changing geometry.
\cleardoublepage
{%
\begin{flushright}
  \null\vspace{\stretch{1}}
  To the memory of my mother Rosalia (1946--2016)\\
  ~\\
  Alla memoria di mia madre Rosalia (1946--2016)

  \vspace{\stretch{3}}
  \null
\end{flushright}}