### 2017-08-17 TitlePage.tex
TitlePage.tex: refactor (enlarge scope of some declarations)

{
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}
\large
\sffamily
% https://tex.stackexchange.com/a/241121/1340
\linespread{1.5}
\selectfont
\begin{titlepage}
% Declarations, in scope until \end{titlepage}.
\makeatletter % for \@title and \@author
\centering
%
% Huge titles are broken by cm-super. Using lmodern fixes it.
\selectfont

\vspace*{-2cm}

{\huge \bfseries
  \@title%
  \par% Must be in scope of the \huge declaration to get correct spacing.
}

\vspace{4cm}

{\bfseries Dissertation}\\
der Mathematisch- und Naturwissenschaftlichen Fakult\"at\\
der Eberhard Karls Universit\"at T\"ubingen\\
zur Erlangung des Grades eines\\
Doktors der Naturwissenschaften\\
(Dr.\ rer.\ nat.)

\vspace{4 cm}

vorgelegt von \\
\@author \\
aus Catania (Italien)

\vspace{2 cm}

T\"ubingen\\
2017

\vfill
\end{titlepage}
\clearpage
% TODO: insert here page 2 with the needed thesis boilerplate.
%
} % End of group with special font declarations.
%
\cleardoublepage
{%
\begin{flushright}
  \null\vspace{\stretch{1}}
  To the memory of my mother Rosalia\\
  ~\\
  Alla memoria di mia madre Rosalia

  \vspace{\stretch{3}}
  \null
\end{flushright}}