### 2017-08-24 TitlePage.tex
Revise titlepage layout
* Ensure it gets the right pagestyle
* Move \cleardoublepage to the scope of the empty pagestyle, so that the
  produced pages use it.
* Use `\vspace[*]` and `\stretch` (that is, `fill`) carefully, instead of other
  mixed vertical skips.
* Increase title `\linespread` to 1.3.

{
\pagestyle{empty}
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}
\large
\sffamily
% https://tex.stackexchange.com/a/241121/1340
\linespread{1.5}
\selectfont
\begin{titlepage}
% Declarations, in scope until \end{titlepage}.
\makeatletter % for \@title and \@author
\centering
%
% Huge titles are broken by cm-super. Using lmodern fixes it.
\selectfont

\vspace*{\stretch{-2}}

{\linespread{1.3} \bfseries \Huge \selectfont
  \@title%
  \par% Must be in scope of the \huge declaration to get correct spacing.
}

\vspace{\stretch{4}}

{\bfseries Dissertation}\\
der Mathematisch- und Naturwissenschaftlichen Fakult\"at\\
der Eberhard Karls Universit\"at T\"ubingen\\
zur Erlangung des Grades eines\\
Doktors der Naturwissenschaften\\
(Dr.\ rer.\ nat.)

\vspace{\stretch{4}}

vorgelegt von \\
\@author \\
aus Catania (Italien)

\vspace{\stretch{2}}

T\"ubingen\\
2018

\vspace{\stretch{-1}}
\end{titlepage}
\clearpage
%
\vspace*{\fill}

Gedruckt mit Genehmigung der Mathematisch-Naturwissenschaftlichen Fakultät der
Eberhard Karls Universität Tübingen.

% 3 bigskips in a row.
% https://tex.stackexchange.com/a/245663/1340
\vspace{\dimexpr \bigskipamount * 3 \relax}
%
% https://tex.stackexchange.com/a/16261/1340
\begin{tabular}{@{}ll@{}}
Tag der mündlichen Qualifikation: & 26.01.2018\\
Dekan:               & Prof.\ Dr.\ Wolfgang Rosenstiel\\
1.\ Berichterstatter: & Prof.\ Dr.\ Klaus Ostermann\\
2.\ Berichterstatter: & Prof.\ Dr.\ Fritz Henglein
\end{tabular}
%
\vspace{-1cm} %XXX revise when changing geometry.
%
%
\cleardoublepage
{% Group with normal linespread
\linespread{1}
\selectfont
\begin{flushright}
  \vspace*{\stretch{1}}
  To the memory of my mother Rosalia (1946--2016)\\
  ~\\
  Alla memoria di mia madre Rosalia (1946--2016)

  \vspace{\stretch{3}}
\end{flushright}
} % End group with normal linespread
\cleardoublepage % End pages in the empty pagestyle.
} % End of group with special font declarations.