### 2019-12-31 thesis-main.tex
thesis-main.tex: Update includeonly statements

%&thesis-main
% The above line loads a precompiled format coming from precompiling thesis-main.ltx.

\usepackage{fancyhdr-setup} % Mine

% Ensure that ligatures in the output can be copy-pasted correctly.
\pdfgentounicode=1
% Loading this file in the format has no effect.
\input{glyphtounicode}

% For some reason this breaks in the static preamble, or at the point it's defined, not sure.
\newunicodechar{⊕}{\texorpdfstring{\ensuremath{\oplus}}{⊕}}
\newunicodechar{Δ}{\texorpdfstring{\ensuremath{\Delta}}{Δ}}
\newunicodechar{τ}{\texorpdfstring{\ensuremath{\tau}}{τ}}
\newunicodechar{→}{\texorpdfstring{\ensuremath{\rightarrow}}{→}}
% Predefined by syntaxmacros.sty
%\newunicodechar{λ}{\texorpdfstring{\ensuremath{\lambda}}{λ}}

% Height of an empty line in lhs2TeX output
% The default value is 0.66084ex. XXX don't adjust it now.
%\setlength{\blanklineskip}{1ex}

% Toggles full and poplAsChapter are initialized to false.

% For adapting to compiling the whole thesis, or only parts.
\toggletrue{full}
\toggletrue{tocs}

\toggletrue{poplAsChapter}
\iftoggle{full}{
\includecomment{fullCompile}
\excludecomment{partCompile}
\toggletrue{tocs}
}{% else
\excludecomment{fullCompile}
\includecomment{partCompile}
}

% For developing tocs
%\toggletrue{tocs}

\iftoggle{full}{
}{%else
% Only for development.
%
% Requires that other files are compiled with include, and can only be used in
% incremental runs: a first run has to be made without includeonly.
% Also, include can add extra page breaks, so maybe switch back to input at the end.
% From http://web.science.mq.edu.au/~rdale/resources/writingnotes/latexstruct.html.
%
%\includeonly{TitlePage,Abstract}
\includeonly{TitlePage,Abstract,Abstract-de,chap-intro}
%\includeonly{Abstract,biblio,acks}
%
%\includeonly{TitlePage}
%\includeonly{aosd13/graphs/evalResLos,aosd13/paperBody}
%\includeonly{chap-intro-incr,chap-diff-examples,chap-diff-correct-formal,chap-chs}
%\includeonly{chap-eq-reason,chap-th-extensions}
%\includeonly{pldi14/sec-practice}
%\includeonly{popl18-sections}
%\includeonly{popl18/cts-intro,popl18/cts-motivation,popl18/formalization,popl18/cts-case-studies,popl18/cts-relwork,popl18/cts-conclusion}
%\includeonly{chap-towards-sysf,pldi14/sec-rw,chap-incr-concl}
%\includeonly{pldi14/sec-preliminaries,backmatter,chap-operationally,defunc}
%\includeonly{acks,biblio}
}
% Build tikz pictures into separate files. Relies on setup in static preamble.
\tikzexternalize
\begin{document}

\frontmatter
\include{TitlePage}

% Avoid having the word "Abstract" in the header of next page.
\pagestyle{plain}
\include{Abstract}
\include{Abstract-de}

\cleardoublepage
\pagestyle{custom}

\iftoggle{tocs}{
\tableofcontents
\cleardoublepage
\listoffigures
\cleardoublepage
\listoftables
}{}

\mainmatter

\include{chap-intro}

% Test case for issue with →.
%\part{Optimizing Collection Queries by Reification${λ_{\ensuremath{\rightarrow}→}→}$}

% WARNING: contains part command
\include{aosd13/paperBody}

% WARNING: contains part command
\include{chap-intro-incr}
\include{chap-diff-examples}
\include{chap-diff-correct-formal}
\include{chap-chs}
\include{chap-eq-reason}
\include{chap-th-extensions}

\include{pldi14/sec-practice}

\iftoggle{poplAsChapter}{
\include{popl18-sections}
}{
\newcommand{\poplPath}[1]{popl18/#1}
\chapter{Cache-transfer-style conversion}
\label{ch:cts}
\include{popl18/cts-intro}
\include{popl18/cts-motivation}
\include{popl18/formalization}
\include{popl18/cts-case-studies}
\include{popl18/cts-relwork}
\include{popl18/cts-conclusion}
}

\include{chap-towards-sysf}

% Merge relworks?
\include{pldi14/sec-rw}

%\pg{DROP except conclusions}
%\include{new-stuff}

% Dropped
%\include{pldi14/sec-concl}
\include{chap-incr-concl}

\appendix

% WARNING: contains part command
\include{pldi14/sec-preliminaries}
\include{backmatter}
\include{chap-operationally}
\include{defunc}

\backmatter
\include{acks}

% Use include on the \bibliography line avoids always typesetting the bibliography.
\include{biblio}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: