### 2020-01-20 acks.tex
First round

% Emacs, this is -*- latex -*-!

\chapter{Acknowledgments}
Writing a PhD thesis is a significant challenge, both professionally and
personally; in a sense, it's a solitary challenge, but in fact I've had so much
help from so many, both in the research and in the personal growth that led me
here. In fact, this page might have been the hardest to write.

I'll start with my professional acknowledgments. I want to thank my supervisor
Klaus Ostermann for encouraging me to do a PhD and for his years of patient and
kind supervision; thanks also to Fritz Henglein for agreeing to review my
thesis. Christian Kästner helped me become a better researcher and writer.
Tillmann Rendel and Sebastian Erdweg sharpened my ability to articulate my
arguments.

Many colleagues helped directly with work presented in this thesis. In
particular, I'd like to thank again my colleagues Cai Yufei, Tillmann, Sebastian
and Philipp Schuster, and my collaborators Lourdes Del Carmen Gonzàlez Huesca and
Yann Règis-Gianas.

% This page was
% Writing a PhD thesis is a significant step in one's life; hence, before thanking
% those who helped my work directly, I want to thank those who helped me become
% the person I am today.

% % From PLDI14 (?)
% % From ILC17
% I thank Cai Yufei, Tillmann Rendel, Lourdes Del Carmen Gonz\`alez Huesca, Yann
% R\`egis-Gianas, Philipp Schuster, Sebastian Erdweg, Robert Atkey, Gabriel
% Scherer and Marc Lasson for helpful discussions on this project.

% Pasted again from papers:

% - AOSD13
% I thank Sebastian Erdweg for helpful discussions on
% this project, Katharina Haselhorst for help
% implementing the code generator, and the anonymous reviewers, Jacques Carette and Karl Klose
% for their helpful comments on this chapter.

% - PLDI14:
% We would like to thank Ingo Maier, Erik Ernst,
% Tiark Rompf, Sebastian Erdweg, Christoph Koch and the PLDI '14

% - ESOP19
% \paragraph{Acknowledgments} We are grateful to anonymous reviewers:
% they made important suggestions to help us improve our technical
% presentation.
% We also thank
% Lourdes del Carmen González Huesca,
% Klaus Ostermann, Sebastian Erdweg for helpful discussions on
% this project.
% This work was partially supported by DFG project 282458149
% and by SNF grant No. 200021\_166154.